### 2019-02-01 cse517/proposal/main.tex
Checking in paper stuff and updated gitignore to master. (#2)
* paper writing

* Starting 517 proposal.

* Woops we don't want all of it...

* Checking in files we care about and new gitignore.

%
% File acl2018.tex
%
%% Based on the style files for ACL-2017, with some changes, which were, in turn,
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2018}
\usepackage{times}
\usepackage{latexsym}

\usepackage{url}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{CSE 517: Project Proposal}

\author{
  Deric Pang \quad
  Joshua Bean \\
  Paul G. Allen School of Computer Science \& Engineering \\
  University of Washington \\
  Seattle, WA, USA \\
  {\tt \{dericp, jbean96\}@cs.washington.edu} \\
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

\section{Introduction}

We propose improving natural language inference models by incorporating
syntactic or contextual features into existing models. At its time of
publication, the well-known decomposable attention neural network by
\citet{parikh2016decomposable} obtained state of the art results on the
SNLI dataset (Bowman et al.
2015) while using drastically fewer parameters than previous NLI models. We
want to extend this model and evaluate it on SNLI (Bowman et al. 2015), SNLI
hard (Gururangan et al. 2018), MultiNLI (Williams et al. 2017), MultiNLI hard
(Gururangan et al. 2018), and possibly even domain specific datasets like
Scitail (Khot et al. 2018).

\bibliographystyle{acl_natbib}
\bibliography{main}

\end{document}