### 2019-02-01 cse517/proposal/main.tex
First draft of project proposal.

%
% File acl2018.tex
%
%% Based on the style files for ACL-2017, with some changes, which were, in turn,
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}

\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{CSE 517: Project Proposal}

\author{
  Deric Pang \quad
  Joshua Bean \\
  Paul G. Allen School of Computer Science \& Engineering \\
  University of Washington \\
  Seattle, WA, USA \\
  {\tt \{dericp, jbean96\}@cs.washington.edu} \\
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

\section{Introduction}

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts.
We seek to improve NLI models by adding syntactic
features into existing models. The well-known decomposable attention (DA)
neural network \citep{Parikh2016-em} obtained state of the art results on the
SNLI \citep{Bowman2015-is} dataset at its time of publication while using
drastically fewer parameters than previous NLI models.

We intend to extend this
model and evaluate it on SNLI, MultiNLI \citep{Williams2017-uh}, the SNLI and
MultiNLI hard splits \citep{Gururangan2018-lj}, and smaller domain specific
datasets like Scitail \citep{Khot2018-th}.

\section{Previous Work}

Our project will build on previous work done by \citet{Pang2018-syntail}.

In previous work, we created the syntail-v1 model whose architecture is
shown in figure \ref{figure:v1}.
We have obtained mixed results as shown in table \ref{table:v1-accuracies} when
naively incorporating syntactic information
into the DA model.

Wihthout ELMo, syntail-v1 improves an impressive 8.4\% in test accuracy over
the baseline decomposable attention model on the SciTail dataset. However,
once ELMo is added to the model, almost the exact happens. Adding syntactic
information decreases test accuracy by 6.7\%.

\begin{table}[h]
  \resizebox{0.45\textwidth}{!}{
  \begin{tabular}{c c c c}
    \toprule
    Model & Embedding & Train Acc. & Test Acc. \\
    \midrule
    DA & GLoVe 6b 300d & 89.4\% & 70.1\% \\
    \textbf{Syntail-v1} & \textbf{GLoVe 6b 300d} & \textbf{92.1\%} & \textbf{78.5\%} \\
    \textbf{DA} & \textbf{ELMo} & \textbf{98.4\%} & \textbf{78.0\%} \\
    Syntail-v1 & ELMo & 88.8\% & 71.3\% \\
    \bottomrule
  \end{tabular}}
  \caption{Syntail-v1 and DA model accuracies on the SciTail dataset.}
\label{table:v1-accuracies}
\end{table}

\section{Proposed Methods}
\label{methods}

We plan to use state-of-the-art constituency \citep{Stern2017-co} and
dependency \citep{Dozat2016-gs} parsers to investigate if incorporating
syntactic features improves the performance of the decomposable attention
network. Additionally, we hope to understand how syntactic features interact
with contextual
word embeddings like ELMo \citep{Peters2018-fz} and BERT \citep{Devlin2018-qc}
in a task like NLI.

We will first investigate different ways of incorporating syntactic
information. \citet{Pang2018-syntail} use a constituency parser to obtain an
encoding of the premise and hypothesis that is used when characterizing their
relationship. We believe that incorporating syntactic features earlier in the
pipeline, especially before the premise and hypothesis are attended over one
another, should improve performance.

We will then investigate different model architectures under a multi-task
training objective where the jointly learns to both parse and predict
entailment relationships.

\begin{figure}[h]
  \includegraphics[width=0.45\textwidth]{v1}
  \caption{A simple method of incorporating syntax into the decomposable
    attention model.}
\label{figure:v1}
\end{figure}

\section{Expected Outcomes}

We will compare our models to the DA model as well as syntail-v1. We hope to
evaluate our models on many diverse datasets as stated in section
~\ref{methods}.

We expect that one of our modeling ideas will produce results that improve on
the DA and syntail-v1 models. In the best case, we will show that syntax is
crucial to building an accurate NLI system.

\section{Challenges}

The first challenge involved in our project will be to modify syntail-v1
so that syntactic information is incorporated much earlier. It will also take
significant time to investigate the best parsers to use for our model.
After, we will write a new model under a multi-task training setup, which we
anticipate will be a significant implementation challenge.

We expect to require significant compute power and time to adequately fine-tune
and evaluate our models. Contextual word embeddings are especially costly to
train, and the mixing ratio in a multi-task setup is well-known to be
difficult to tune.

\bibliographystyle{acl_natbib}
\bibliography{main}

\end{document}