### 2019-03-26 paper/main.tex
Cleaning up writeup.

%
% File acl2018.tex
%
%% Based on the style files for ACL-2017, with some changes, which were, in turn,
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{multirow}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\newcommand{\dpcomment}[1]{\textcolor{green}{[#1 --deric]}}
\newcommand{\jbcomment}[1]{\textcolor{orange}{[#1 --josh]}}

\title{Syntactically Informed Natural Language Inference}

\author{
  Deric Pang \quad
  Joshua Bean \\
  Paul G. Allen School of Computer Science \& Engineering \\
  University of Washington \\
  Seattle, WA, USA \\
  {\tt \{dericp, jbean96\}@cs.washington.edu} \\
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  We demonstrate the importance of syntactic information in semantic models by
  extending the decomposable attention model for natural language inference.
  By pipelining the hidden states of a neural syntax parser into the
  decomposable attention model, we achieve greater performance on the SciTail
  dataset than the ESIM model and DA + ELMo.
\end{abstract}

\section{Introduction}

Natural language inference (NLI) is the task of characterizing entailment and
contradiction relationships between texts.  In general, most NLI tasks are
formulated as characterizing the relationship between a pair of sequences---a
premise and a hypothesis. An NLI model should predict whether the hypothesis is
entailed by the premise, contradicts the premise, or is neutral to the premise.

We extend the decomposable attention (DA) model from
\citet{parikh2016decomposable} by incorporating syntactic features.  The DA
model obtained state-of-the-art results on the SNLI \citep{bowman2015large}
dataset at its time of publication while using drastically fewer parameters
than previous NLI models.  We primarily evaluate on the domain-specific SciTail
dataset \citep{Khot2018-th} for both its smaller size and lack of annotation
artifacts when compared to SNLI \citep{gururangan2018annotation}.

Our model, \texttt{syntail}, achieves 4.8\% better test accuracy than the
vanilla DA model and 1.9\% better test accuracy than DA with ELMo
\citep{peters2018deep} all while maintaining the core DA architecture.

\section{Decomposable Attention Model}

The DA model is a simple and easily parallelizable approach to NLI.  We
summarize it here so we can build upon it later.  The model decomposes into
attend, compare, and aggregate steps.

Let $\bm{p} = \langle p_1, \dots, p_{\ell_p} \rangle$ and $\bm{h} = \langle
h_1, \dots, h_{\ell_h} \rangle$ where each $p_i, h_j \in \mathbb{R}^d$ is a
$d$-dimensional word embedding vector.

\paragraph{Attend.} Compute unnormalized attention weights $e_{ij}$ with
a feed-forward neural network $F$:
\begin{align}\label{eq:attend}
    e_{ij} := F(p_i)^\mathsf{T} F(h_j).
\end{align}
Compute the aligned subphrases $\bm{H}i$ and $\bm{P}_j$:
\begin{align}
  \bm{H}_i &:= \sum_{j=1}^{\ell_h}\frac{\exp{(e_{ij})}}
      {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}h_j\,, \notag \\
  \bm{P}_j &:= \sum_{i=1}^{\ell_p}\frac{\exp{(e_{ij})}}
      {\sum_{k=1}^{\ell_h}\exp{(e_{ik})}}p_i\
\end{align}
where $\bm{H}_i$ is the weighted sum over $\bm{h}$ that aligns to $p_i$ and
vice versa for $\bm{P}_j$.

\paragraph{Compare.} Compare each $(p_i, \bm{H}_i)$ and $(h_j, \bm{P_}j)$ pairs
with another feed-forward neural network $G$:
\begin{align}
  \bm{v}^p_i &:= G([p_i, \bm{H}_i])\quad \forall i \in [1,\ldots, \ell_p]\,,
      \notag \\
  \bm{v}^h_j &:= G([h_j, \bm{P}_j])\quad \forall j \in [1,\ldots, \ell_h]\,.
\end{align}

\paragraph{Aggregate.} Aggregate each set of comparison vectors:
\begin{align}
\bm{v}^p = \sum_{i=1}^{\ell_p} \bm{v}^p_i \qquad\,, \qquad
\bm{v}^h = \sum_{j=1}^{\ell_h}  \bm{v}^h_j\,.
\end{align}
and make a prediction with a final feed-forward layer $H$:
\begin{align}\label{eq:predict}
    \hat{\bm{y}} = H([\bm{v}^p, \bm{v}^h]).
\end{align}

\section{Our Model}

The motivation for incorporating syntactic information in an NLI model can be
demonstrated with a simple example. Consider the premise ``Adrian is running
and Alex is swimming." If we present the DA model trained on SNLI with the
hypothesis ``Adrian is swimming," it will confidently predict that this
blatantly incorrect hypothesis is entailed by the premise. Observing any
syntactic parse of the premise will make it obvious that Adrian is in fact
running.

We experiment with features produced by two syntax parsers---the minimal
span-based neural constituency parser \citep{stern2017minimal} and the deep
biaffine attention model for dependency parsing \citep{dozat2016deep}. We also
experiment with different methods of incorporating the syntactic features
extracted from these parsers.

\subsection{Extracting Syntactic Information}

The minimal span-based neural constituency parser and the deep biaffine
attention model for dependency parsing both encode the input sequence with an
LSTM. Using this LSTM, we obtain syntactic features $\bm{s}^p = \langle s^p_1,
\dots, s^p_{\ell_p} \rangle$ and $\bm{s}^h = \langle s^h_1, \dots, s^h_{\ell_h}
\rangle$ from the premise and hypothesis, respectively. An $s_i$ is either the
final hidden state $\bm{r}_i$ of the LSTM at index $i$ or a projected
representation of $\bm{r}_i$.

\subsection{Syntail Architectures}

We will now describe the different model architectures we experimented with.
In general, as the version number increases, the complexity of the model
increases.

\paragraph{v1: Naive late fusion of syntactic features.} The simplest way to
incorporate $\bm{s}$ is to concatenate its final representation to $\bm{v}^p$
and $\bm{v}^h$ as shown in figure~\ref{fig:v1}.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figures/v1.png}
    \caption{Syntail v1 architecture.}
\label{fig:v1}
\end{figure}

The prediction step in equation~\ref{eq:predict} then becomes:
\begin{align}
    \hat{\bm{y}} = H([\bm{v}^p, s^p_{\ell_p}, \bm{v}^h, s^h_{\ell_h}])
\end{align}
where we concatenate the final hidden representation of the premise and
hypothesis when passed through the neural syntax parser with the aggregrated
comparision vectors.  Everything else is identical to the DA model.

\paragraph{v2: Using syntactic features to attend.} We wanted to experiment
with incorporating syntactic features earlier in the model.  Instead of
learning weights on the syntactic features, it is possible to concatenate them
to the projected input representations as show in figure~\ref{fig:v2}.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/v2.png}
  \caption{Syntail v2 architecture.}
\label{fig:v2}
\end{figure}

Computing the unnormalized attention weights in equation~\ref{eq:attend} then
becomes: \begin{align} e_{ij} := [F(p_i), s^p_i]^\mathsf{T} [F(h_j), s^h_i]
\end{align} where instead of calculating the alignment using only the projected
hypothesis and premise vectors, we now also use the final hidden states of the
premise and hypothesis when passed through the neural syntax parser.  This
seems rather naive since this model will not learn any weights on the syntactic
features, but in practice this model performs very well. Everything else is
identical to the DA model.

\paragraph{v3: Using syntactic features to project the premise and hypothesis.}
We decided to try incorporating the syntactic features even earlier in the
model.  Instead of concatenating the final hidden states of the parsed premise
and hypothesis with the projected inputs, we concatenate the final hidden
states of the parsed premise and hypothesis with the inputs immediately before
projecting them. This architecture is shown in figure~\ref{fig:v3}.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/v3.png}
  \caption{Syntail v3 architecture.}
\label{fig:v3}
\end{figure}

Computing the unnormalized attention weights in equation~\ref{eq:attend} then
becomes:
\begin{align}
    e_{ij} := F([p_i, s^p_i])^\mathsf{T} F([h_j, s^h_j]).
\end{align}
Everything else is identical to the DA model.

\paragraph{v4: Projecting the syntactic features before computing attention.}
After noticing a drop in performance from v2 to v3, we hypothesized that it was
due to not allowing our model to transform the syntactic features enough before
calculating attention.  We decided to learn weights to project the final hidden
states of the parsed premise and hypothesis by passing them through an
additional fully-connected layer before concatenating them to the inputs. This
architecture is shown in figure~\ref{fig:v4}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{figures/v4.png}
    \caption{Syntail v4 architecture.}
\label{fig:v4}
\end{figure}

Computing the unnormalized attention weights in equation~\ref{eq:attend} then
becomes:
\begin{align}
    e_{ij} := F([p_i, J(s^p_i)])^\mathsf{T} F([h_j, J(s^h_j)])
\end{align}
where $J$ is a feed-forward neural network. Everything else is identical to
the DA model.

\subsection{Implementation Details}

\subsubsection{Word Embeddings}

Due to resource constraints, we limited our models to only use 300-dimensional
6B GloVe word embeddings \citep{pennington2014glove}.  Additionally, contextual
embeddings are known to capture syntax in some capacity, so limiting our models
to only use GloVe vectors made it easier to experiment with the syntactic
features in isolation.

\subsubsection{Neural Syntax Parser}

We initially used a constituency parser in syntail. However, after some
experimentation, we found that we were achieving better results with a
dependency parser.  For this reason, we only report results using the deep
biaffine attention model for dependency parsing.  Dependency parsers focus on
encoding the the relationships between words as opposed to determining
constituents.  Intuitively, if we consider our example sentence, ``Adrian is
running and Alex is swimming,'' the dependency arcs will be more informative of
who is doing what.

\subsection{Hyper-parameters}

We use 200-dimensional projections of the word embeddings. Our feed-forward
neural networks are 2 layered and have hidden dimensions of 200. We train with
both Adam \citep{kingma2014adam} and Adagrad \citep{duchi2011adaptive} and
report the better result.

\section{Experiments}

We compare our model's performance to the decomposable attention model
\citep{parikh2016decomposable} and the ESIM model \citep{chen2016enhanced}.
Early on in our experiments, we noticed that features from a dependency parser
seemed to perform better than those from a constituency parser.  Therefore, all
of our models use the deep biaffine attention model for dependency parsing
\citep{dozat2016deep}.

\paragraph{Datasets.} For practical reasons, we only experiment on the SciTail
dataset. SciTail (27k pairs) is significantly smaller than SNLI (570k pairs)
\citep{bowman2015large} or MultiNLI (433k pairs) \citep{williams2017broad}
which makes it feasible to try many different architectures given limited
resources.

Additionally, \citet{gururangan2018annotation} showed that SNLI and MultiNLI
have significant annotation artifacts whereas SciTail is the first entailment
set to be completely derived from sentences that exist ``in the wild.''

\paragraph{Baselines.}
We list baseline performance reported by \citet{Khot2018-th} of a majority
class predictor, n-gram model, and ESIM \citep{chen2016enhanced} in
table~\ref{table:experiments}. We reimplemented and retrained the DA model and
obtained similar test accuracies as reported by \citet{chen2016enhanced}.
Adding ELMo to the DA model improved test accuracy performance by approximately
3\%.

\subsection{Results}

\begin{table}[htb]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{cccccc} 
\toprule
Model          & Embeddings      & Dev. Acc.     & Test Acc.     \\ 
\midrule
Majority Class & n/a             & 63.3          & 60.3          \\
N-Gram         & n/a             & 65.0          & 70.6          \\
ESIM           & GloVe 840B 300d & 70.5          & 70.6          \\
DA             & GloVe 6B 300d   & 70.4          & 72.6          \\
DA             & ELMo            & 79.1          & 75.5          \\
\midrule
Syntail v1     & GloVe 6B 300d   & 78.4          & \textbf{77.4} \\
Syntail v2     & GloVe 6B 300d   & \textbf{82.3} & \textbf{77.4} \\
Syntail v3     & GloVe 6B 300d   & 79.0          & 76.1          \\
Syntail v4     & GloVe 6B 300d   & 81.3          & 77.0          \\
\bottomrule
\end{tabular}}
\caption{Dev and test accuracies on SciTail.}
\label{table:experiments}
\end{table}

Our experimental results can be found in table~\ref{table:experiments}. All of
our syntail models achieve better test accuracy than the baseline models. This
demonstrates that the syntax parser is providing features that are both useful
for determining entailment relationships and difficult for the baseline models
to learn on their own.

We were surprised that the simpler v1 and v2 models outperformed the more
complicated v3 and v4 models. This is likely due to the syntactic information
being fused later in v1 and v2. The later the features are fused, the closer
they are to the loss function and the more the model can learn to use them.

It is also interesting that adding and additional transformation on the
syntactic features in v4 decreases performance.  Again, this is likely due to
the early fusion problem.  The additional power of the model is overwhelmed by
the detrimental effects of the vanishing gradient problem.  It is also possible
that the randomly initialized weights of the feed-forward neural network
struggle to learn how to summarize the syntactic features.

One downside of using a pretrained syntax parser is that the data SciTail is
created from (science exam questions) is in a very different domain than the
data used to create the Penn Treebank (news articles)
\citep{marcus1993building}.  This likely significantly reduces the accuracy of
the syntax parser on SciTail. Retraining the syntax parser on in-domain data
should produce better syntactic features and further improve the performance of
our model.

\subsection{Analysis of Syntactic Features}

We discuss a specific in-domain input example shown in
table~\ref{table:example} to better understand the effects of syntactic
features. The predictions on this input made by the DA and syntail v1 models
are shown in table~\ref{table:pred-example}.

\begin{table*}[t]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabular}{c c}
    \toprule
    Premise & Hypothesis \\
    \midrule
    The continental crust is less dense than the mantle but the oceanic crust
        is not. & The oceanic crust is more dense than the mantle. \\
    \bottomrule
  \end{tabular}}
  \caption{A pair of sentences where syntail performs particularly well. The
      hypothesis is entailed by the premise.}
\label{table:example}
\end{table*}

\begin{table}[htb]
  \centering
  \begin{tabular}{c c c}
    \toprule
    Model & Entails (correct) \\
    \midrule
    DA & 29.5\% \\
    Syntail v1 & \textbf{67.9\%} \\
    \bottomrule
  \end{tabular}
  \caption{Entailment probabilities generated by the DA and syntail v1 models
      on the input from table~\ref{table:example}.}
\label{table:pred-example}
\end{table}

Syntail v1 correctly determines the relationship of this pair of sentences
while the DA model does not.  To investigate why, we visualize the attention
weights in figures~\ref{fig:da-attention-weights} and
\ref{fig:sytail-attention-weights}.  We can see that the attention produced by
the DA model is nonsensical and focuses solely on ``is.'' It seems that by
incorporating syntactic information, syntail v1 is forced to learn a more
reliable attention function.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/da-h2p-attention.png}
  \caption{Hypothesis to premise attention weights of the DA model when
      processing the input from table~\ref{table:example}.}
\label{fig:da-attention-weights}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/syntail-h2p-attention.png}
  \caption{Hypothesis to premise attention weights of syntail v1 when
      processing the input from table~\ref{table:example}.}
\label{fig:sytail-attention-weights}
\end{figure}

\section{Related Work}

Previous works have experimented with incorporating syntactic features for
semantic tasks.  \citet{chen2016enhanced} use tree structures generated by a
constituency parser to improve their ESIM model. They encode the
phrase-structure trees with a Tree-LSTM and use its hidden states.
\citet{Khot2018-th} incorporate syntactic features by proposing a model
architecture which can use any graph-based syntactic/semantic structure.

Other works have experimented with incorporating syntactic learning objectives.
\citet{strubell2018linguistically} achieved state-of-the-art performance for a
model using predicted predicates and standard word embeddings on CoNLL-2005 SRL
by performing multi-task training on additional syntactic tasks.
\citet{swayamdipta2017frame} use hard-coded constituency and dependency parse
features and a syntactic scaffold to extend the softmax margin segmental RNN
(segRNN) model.

\section{Conclusion}

By making simple extensions to the DA model, our syntail models outperform ESIM
and DA + ELMo model on SciTail.  Our results demonstrate the importance of
syntactic information in semantic models and motivate future research into
syntactically informed models.

\bibliographystyle{acl_natbib_nourl}
\bibliography{references}

\end{document}